---
title: Examples
author: Jürgen Wilbert
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#library(tidyverse)
#library(scan)
library(scplot)
```

# scplot

We started developing a new add-on package to `scan`for visualizing single-case data: `scplot`. This function will gradually replace the `plot.scdf()` (short: `plot()`) function already included in `scan` and finally be included into to the `scan` package. 
Here are some advantages of using `scplot` over the standard scan `plot` function:

- `scplot` is already much more versatile than `plot` has been. 
- `scplot` was designed to encompass a pipe style coding which is much cleaner, more intelligible and easier to code. 
- `scplot` is based on `ggplot2` and produces a `ggplot2` object which can be modified and extended to any wishes.

We consider the state of `scplot` to be *experimental*. That is, the code and syntax might change in future versions so backward compatibility is not guaranteed.

But we will keep the "old" `plot.scdf` in future versions of `scan`.

Here are a few plots that have been generated with `scplot` to demonstrate its possibilities.

## Install `scplot`

scplot is hosted as a gitHub project at https://github.com/jazznbass/scplot.
You can install it with `devtools::install_github("jazznbass/scplot", dependencies = TRUE)` from your R console.
Make sure you have the package `devtools` installed before.
The package has to be compiled. When you are running R on a Windows machine you also have to install Rtools. Rtools is not an R package and can be downloaded from CRAN at https://cran.r-project.org/bin/windows/Rtools/. MacOs and Linux users usually do not need to take this extra step.

## The standard style

```{r}
scplot(exampleAB)
```

## Add statlines

```{r}
scplot(exampleABC) %>%
  add_statline("mean", color = "darkred")
```

## Add a legend

```{r}
scplot(exampleABC) %>%
  add_statline("mean", color = "darkred") %>%
  add_statline("min", phase = "B", width = 0.2, color = "darkblue") %>%
  add_legend()
```

## A simple style with axis settings

```{r}
scplot(exampleA1B1A2B2) %>% 
  set_xaxis(increment_from = 0, increment = 5) %>%
  set_yaxis(limits = c(0, 50)) %>%
  set_ylabel(angle = 0) %>%
  add_theme("default", "small") %>% 
  set_casenames(type = "strip") %>%
  set_panel(fill = c("grey90", "grey80"))
```

## Add labels

```{r}
scplot(exampleABC) %>% 
  set_xaxis(increment = 2) %>%
  set_yaxis(color = NA) %>%
  add_labels(text = list(color = "black", size = 0.7), 
             background = list(fill = "grey98"), nudge_y = 7) %>%
  add_theme("minimal")
 
```

## Rotated axis values, removed dots, and some adjustments on phasenames

```{r}
scplot(example_A24) %>% 
  add_theme("default") %>%
  add_statline("lowess", color = "darkred", width = 1.5) %>%
  add_statline("loess", color = "red", width = 1.5) %>%
  add_statline("movingMean", color = "lightpink", width = 1.5) %>%
  set_xaxis(size = 0.8, angle = 35) %>%
  set_dataline(size = 0) %>%
  add_legend(position = c(0.8, 0.75), background = list(color = "grey50")) %>%
  set_phasenames(c("no speedlimit", "with speedlimit"), position = "left", 
                 hjust = 0, vjust = 1) %>%
  set_casenames("") %>%
  add_title("Effect of a speedlimit on the A24") %>%
  add_caption("Note: Moving mean calculated with lag two.", face = 3) %>%
  add_ridge(color = "lightblue")
  
```

## Add a ridge

```{r}
scplot(exampleA1B1A2B2_zvt) %>% 
  add_ridge("yellow")
```

## Mulitple variables in a plot with an added legend

```{r multi-variables}
scplot(exampleAB_add) %>%
  add_dataline("cigarrets", color = "darkblue", size = 1) %>%
  add_statline("trend", linetype = "dashed") %>%
  add_statline("mean", variable = "cigarrets", color = "darkred") %>%
  add_marks(positions = c(14,20), size = 3, variable = "cigarrets")%>%
  add_marks(positions = "cigarrets > quantile(cigarrets, 0.75)", size = 3) %>%
  set_xaxis(increment = 5) %>%
  set_phasenames(color = NA) %>%
  set_casenames(type = "strip") %>%
  add_legend(title = list(color = NA), text = list(face = 3)) %>%
  set_panel(fill = c("lightblue", "grey80")) %>%
  add_ridge(color = "snow", variable = "cigarrets") %>%
  add_labels(variable = "cigarrets", nudge_y = 2, 
             text = list(color = "blue", size = 0.5)) %>%
  add_labels(nudge_y = 2, text = list(color = "black", size = 0.5),
             background = list(fill = "white"))


```

## An entensive example to show-off

```{r}
scplot(exampleA1B1A2B2) %>% 
  set_xaxis(increment = 4, color = "brown") %>%
  set_yaxis(color = "sienna3") %>%
  set_ylabel("Points", color = "sienna3", angle = 0) %>%
  set_xlabel("Weeks", size = 1, color = "brown") %>%
  add_title("Points by week", color = "sienna4", face = 3) %>%
  add_caption("Note: An extensive Example.",
              color = "black", size = 1, face = 3) %>%
  set_phasenames(c("Baseline", "Intervention", "Fall-Back", "Intervention_2"), 
                 size = 0) %>%
  add_ridge(alpha("lightblue", 0.5)) %>%
  set_casenames(labels = sample_names(3), color = "steelblue4", size = 0.7) %>%
  set_panel(fill = c("grey80", "grey95"), color = "sienna4") %>%
  add_grid(color = "grey85", size = 0.5) %>%
  set_dataline(color = "black", width = 0.5, linetype = "solid", 
               dots = "sienna4", size = 0.5, shape = 18) %>%
  add_labels(text = list(color = "sienna", size = 0.7), nudge_y = 4) %>%
  set_seperator(extent = 0.8, size = 0.5, 
                linetype = "solid", color = "sienna") %>%
  add_statline(stat = "trendA", color = "tomato2") %>%
  add_statline(stat = "max", phase = c(1, 3), linetype = "dashed") %>%
  add_marks(case = 1:2, positions = 14, color = "red3", size = 2, shape = 4) %>%
  add_marks(case = "all", positions = "values < quantile(values, 0.1)", 
            color = "blue3", size = 1.5) %>%
  add_marks(positions = outlier(exampleABAB), color = "brown", size = 2) %>%
  add_text(case = 1, x = 5, y = 35, label = "Interesting", 
           color = "darkgreen", angle = 20, size = 0.7) %>%
  add_arrow(case = 1, 5, 30, 5, 22, color = "steelblue") %>%
  set_background(fill = "white") %>%
  add_legend()

```
